Decision-Optimal Probabilistic Forecast Reconciliation for Expected Shortfall in Hierarchical Time Series

### 2025-08-19 phd_application/phd_propossal.Rmd
Update phd_propossal.Rmd

Existing reconciliation methods for point forecasts can be generalised to reconcile probabilistic forecasts. However, these approaches focus on point forecasts and are optimised for scoring rules such as mean squared error (MSE). There is a gap in the literature on how *optimal* this approach is with respect to the downstream decision. As the current expansion of point reconciliation methods no longer guarantee optimality for probabilistic forecasts, such as quantiles or expected shortfall (ES), a more principled approach is needed. This leaves the question: *"How can probabilistic reconciliation be optimised with respect to different loss functions targeted at the downstream decision?"*

Among the risk measures, quantiles, also known as Value-at-Risk (VaR), is a common choice, but it has limitations. VaR ignores the severity of losses beyond that threshold and is not a coherent risk measure, since it does not satisfy the subadditivity property [@Yamai2005-me]. This can lead to inconsistencies in risk assessments across different levels of the hierarchy.

>>>>

Existing reconciliation methods for point forecasts has recently been generalised to reconcile probabilistic forecasts. However, these approaches originally focus on point forecasts and are optimised for scoring rules such as mean squared error (MSE). There is a gap in the literature on how *optimal* this approach is with respect to the downstream decision. The current expansion of point reconciliation methods no longer guarantee optimality for probabilistic forecasts, such as quantiles or expected shortfall (ES). This leaves the question: *"How can probabilistic reconciliation be optimised with respect to different loss functions targeted at the downstream decision?"*

Although @Panagiotelis2023-fs introduced an score optimisation for reconciliation weights using the energy score or variogram score, they are scoring rules for full multivariate predictive distributions and do not directly address the tail functionals. Among the risk measures, quantiles, also known as Value-at-Risk (VaR), is a common choice, but it has limitations. VaR ignores the severity of losses beyond that threshold and is not a coherent risk measure, since it does not satisfy the subadditivity property [@Yamai2005-me]. This can lead to inconsistencies in risk assessments across different levels of the hierarchy.
